---
title: "Business Analytics Lab Workshop 2"
author: "Kajal Chokshi"
date: "June 5th 2017"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
subtitle: CME Group Foundation Business Analytics Lab
---

###About

Today the lab focuses on data outliers, data preparation, and data modeling.
This lab requires the use of Microsoft Excel, R, and erdplus

###SetUp

Make sure to download the folder titled 'bsad_lab2' zip folder and extract the
folder to unzip it. Next, we must set this folder as the working directory. The
way to do this is to open R Studio, go to 'Session', scroll down to 'Set Working
Directory', and click 'To Source File Location'. Now, follow the worksheet
directions to complete the lab.

---------

###Task 1

First, we must calculate the mean, standard deviation, maximum, and minimum for
the Age column using both R and Excel.

In R, we must read in the file again, extract the column and find the values that 
are asked for.

Average Age: 
```{r}
mydata <- read.csv(file="data/creditrisk.csv") 
age <-mydata$Age 
mean(age)
```

Standard Deviation of Age:
```{r}
sd(age)
```

Maximum of Age: 
```{r}
max(age)
```

Minimum of Age: 
```{r}
min(age)
```

Next, compute the same thing using Excel and compare results. Post a screenshot
of this using the directions found in Worksheet 1. To screenshot, you can use
the snipping tool and save the PNG image to the same folder all of the other
data is.

Next, use the formula from class to detect any outliers. Right the code to find
the IQR by hand, and check it with the function that is given in R.

```{r} 
lowerq <- quantile(age)[2] 
upperq <- quantile(age)[4] 
iqr <- IQR(age) 
iqr
```

The formuala below calculates the threshold. 
```{r} 
upperthreshold <- (iqr * 1.5) + upperq 
lowerthreshold <- lowerq - (iqr * 1.5)
upperthreshold 
lowerthreshold

```

Are there any outliers? How many? It can also be useful to visualize the data
using a box and whisker plot. The boxplot below supports the IQR we found of 15
and upper and lower threshold.

```{r} 
boxplot(age) 
```

---------------

###Task 2

Next, we must calculate the mean for checking in Excel. Take a screenshot of
this process and paste it here. How are the missing values are treated in the
Excel calculation?

Now, read the file into R. 

```{r}
newdata <-read.csv(file="data/creditriskorg.csv")
newdata 
```

We observe that the column names are shifted down below. So, we must make sure
to use the command skip and set the header to true.

```{r} 
newdata <- read.csv(file="data/creditriskorg.csv",
skip=1,header=TRUE,sep=",") 
newdata 
```

To calculate the mean for Checking in R, follow Worksheet 1. Extract the
Checking column first and then find the average using the function built in R.
What happens when we try to use the function?

```{r} 
checking <- newdata$Checking 
```

To resolve the error, we must remove understand where it is coming from. There
are missing values in the csv file, which is quite common as most datasets are
not perfect. Additionally, there are commas within the excel spreadsheet, and R
does not recognize that '1,234' is equivalent to '1234'. Lastly, there are '$'
symbols throughout the file which is not a numerica symbol either.

The sub function replaces these symbols with something else. So, in order to
remove the comma in the number "1,234", we must substitute it with just an empty
space.

```{r} 
checking <- sub(",", "", checking) #substitute comma with blank in all of checking 
checking <- sub("\\$", "", checking) #substitute dollar sign with blank in all of checking 
checking <-  as.numeric(checking) #numeric convert 
mean(checking, na.rm=TRUE) #mean with NA removed 
```

What are some other ways to clean this data in R? How about in Excel? 

-------------

###Task 3

Now, we will look at divvybike data. Go to ```https://www.divvybikes.com/data?```Download latest 2016 Q3/Q4 data (zip file) and check the README file. 

Open the file Divvy_Trips_2016_Q4 in Excel. Per the Data License Agreement who owns the data, and can the data be correlated with other sources?

Note the size of the file, the number of columns and of rows here.??? Identify the unique entities and fields.

Define a relational business logic for the column field 'tripduration'.

Using www.erdplus.com draw a star schema using the following three tables:- A Fact table for Trip- A Dimension table for Station- A Dimension table for User

