---
title: "Business Analytics Lab Workshop 4"
author: "Kajal Chokshi"
date: "June 7th 2017"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
subtitle: CME Group Foundation Business Analytics Lab
---

###About

R can be used to make basic visual analytics, which can be helpful in understanding the data holistically. Additionally, R can help find correlations between variables and create scatter plots. 

In this lab we will delve deeper into the marketing data set with R visual analytics and use Tableau to visualize the credit risk data set. 

###SetUp

Make sure to download the folder titled 'bsad_lab4' zip folder and extract the folder to unzip it. Next, we must set this folder as the working directory. The way to do this is to open R Studio, go to 'Session', scroll down to 'Set Working
Directory', and click 'To Source File Location'. Now, follow the worksheet directions to complete the lab.

----------

###Task 1

First, read in the 'marketing.csv' file and make sure all the columns are capture. 

```{r}
mydata <-read.csv(file="data/marketing.csv")
mydata
```

Now, we will use somme visuals to observe some differences. 

First create a bar chart of the sales variable. Remember to extract the sales variable before calling it.  

```{r}
sales <- mydata$sales
barplot(sales)

```

Next, create a histogram of sales. 

```{r}
hist(sales)
```

What are some observable differences between the boxplot and the histogram? Please record your observations. 

Lastly, create a pie chart of sales.

```{r}
pie(sales)
```

Why is this visual not a good representation of the data? What makes the histogram a better representation?

----------

###Task 2

The previous task focused on visualizing one variable. A good way to visualize two variables would be a scatter plot or a correlation matrix. 

Lets create a scatterplot of Sales vs. Radio. 

```{r}
radio <- mydata$radio
plot(radio,sales)
```

It might be easier to notice a trend if there was a line that fit through the points. So, lets repeat using the function scatter.smooth, which adds a smooth curve computed by loess, a method to locally smooth a curve. 

```{r}
scatter.smooth(radio,sales)
```

Now, repeat below for Sales vs. POS, following the example that was just provided. Observe and note any trends in the scatter plots and potential relationships below. 

To quantify the data, its best to look at the correlation between two variables.

```{r}
cor(sales,radio)
```

What does this correlation explain? 

Below, compute the correlation between every pair of variables (Sales & TV, Sales & POS, Sales & Pape). Make sure to extract the variables first and then compute the correlation. 

----------

###Task 3

Follow the directions on the worksheet.  

1. Refer to file 'creditrisk.csv' from worksheet 1. 

2. Start Tableau and enter your LUC email if prompted.

3. Import the file into Tableau. Choose the Text File option when importing

4. In a new sheet tab drag-drop 'Age' to Rows and select the histogram view. Note the bar like view and breakdowns. Which age 'bin' has the highest count? What is the count?

5. Drag-drop 'Marital Status' to the Marks Color. Which age 'bin' has highest divorce count? What is the count?

6. From 'Show Me' menu change the view to bubbles, and treemaps.

7. Explain how age and marital status data are represented in each of the three views (histogram, bubbles, & treemaps).

8. If you were to show the age with the highest divorce count which of the three views would you select and why?

9. In a new sheet tab drag-drop Months Employed to Rows and Age to Columns. Switch the type for each measure from Sum to Dimension. You should obtain what looks like a scatter plot.

10. Add to the scatter plot the Gender by drag-drop to the Marks Color. Share observations

11. In a new sheet generate a view of your own creation and share your insights

Post screenshots of all your work from Tableau below. Make sure to include the comments if prompted. 































































