---
title: "Business Analytics Lab Workshop 6"
author: "Kajal Chokshi"
date: "June 7th 2017"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
subtitle: CME Group Foundation Business Analytics Lab
---

###About

In this lab we will focus on linear and non-linear programming. Linear programming, as discussed in the previous lab, works with simple and multiple linear regression techniques; sometimes the variables have completely direct or completely non-direct relationships and these techniques can model them. Sometimes, however, the variables do not predict each other in a linear way. For example, looking at the stock market vs. time, we know that generally the market was booming before the crash, then the market crashed and the great depression hit, and slowly the market started to rise again. This pattern is not linear, and in fact a non-linear programming technique can be used to model it and predict the value of the market based on the year. 

In this lab, we will explore topics like optimization, solve a marketing model, and perform linear and non-linear regression on the cost of servers.

###SetUp

Make sure to download the folder titled 'bsad_lab6' zip folder and extract the folder to unzip it. Next, we must set this folder as the working directory. The way to do this is to open R Studio, go to 'Session', scroll down to 'Set Working Directory', and click 'To Source File Location'. Now, follow the worksheet directions to complete the lab.

----------

###Task 1

In the first part of this lab, we will focus on 

Adding Optimization package to R.  Note the 'l' below are as in 'L'
```{r}
install.packages("lpSolveAPI") #install special package required for the solver
library("lpSolveAPI") # load package library

```


Solving Marketing Model 
```{r}
lprec <- make.lp(0, 2) # model object
lp.control(lprec, sense="max")  # set for maximum
set.objfn(lprec, c(275.691, 48.341))
add.constraint(lprec, c(1, 1), "<=", 350000)
add.constraint(lprec, c(1, 0), ">=", 15000)
add.constraint(lprec, c(0, 1), ">=", 75000)
add.constraint(lprec, c(2, -1), "=", 0)
lprec # view the problem formulation in tabular/matrix form
solve(lprec) # solve 
get.objective(lprec) #display the objective function optimum value
get.variables(lprec) # display the variables optimum values

```


-----------

###Task 2

First, make sure to read the file in. Make sure to extract the two columns to use later. 

```{r}
mydata <- read.csv("data/ServersCost.csv", header=TRUE)
mydata
servers <- mydata$servers
cost <- mydata$cost
```

Next, create a linear model. First, we plot the points to visually see the data. Then create a model as we did in the last lab and plot the model directly on top of the graph.

```{r}
plot(servers,cost)
linear_model <- lm(cost ~ servers)
abline(linear_model, col="blue", lwd=2) # add linear line based on regression model
summary(linear_model)

```

We notice that this model is not very good at predicting the cost. So, we use a transformation and use a nonlinear quadratic model. 

```{r}
servers2 <- servers^2
quad_model <- lm(cost ~ servers+servers2) # of form x + x^2
summary(quad_model)
```

```{r}
predicted2 <- predict(quad_model,data=mydata) # compute the predicted values based on the quad model
par(new=TRUE, xaxt="n", yaxt="n", ann=FALSE) # needed to overlay new points without the labels and annotations
plot(predicted2, col="red") # Use the red color to discern new points
```

Lets try a cubic model just to see how it will work. 

```{r}
servers3<-servers^3
cubic_model<-lm(cost ~ servers+servers2+servers3) # of form x + x^2 + x^3
summary(cubic_model)
predicted3 <- predict(cubic_model,data=mydata) # compute the predicted values based on the cubic model
par(new=TRUE, xaxt="n", yaxt="n", ann=FALSE) # overlay new points without the labels and annotations
plot(predicted3, col="green") # Use the green color to identify new points
```

